---
author: David Carslaw
---

# Polar frequencies 风频图{#sec-polarFreq}

## Background and examples {#sec-polarFreqBack}

我们将简要介绍另一种绘制气象数据中风向和风速分布的方式，和传统的风玫瑰图比较相似，但是在结合污染物浓度变化等方面进行了强化，风频图可以用到所有的数据，考虑到了更多的统计变化，也可以按照不同的时间周期分图展示，如逐年、逐月、一周为单位的逐日变化等等。

这一节用内置的 `mydata` 数据演示风频图使用方法。@fig-polarFreqa 是 `polarFreq` 默认状态下的显示结果。

```{r}
#| label: fig-polarFreqa
#| fig-cap: 使用 `polarFreq` 函数绘制风速和风向。每一个小格子代表一个特定的风向和风速组合，格子的颜色代表出现这一个特定风向和风速的总小时数。颜色所  The number of hours is coded as a colour scale shown to the right. The scale itself is non-linear to help show the overall distribution. The dashed circular grey lines show the wind speed scale. 
#| fig-width: 5
#| fig-height: 5
#| out-width: 60%
# load openair
library(openair)
polarFreq(mydata)
```

By setting `type = "year"`, the frequencies are shown separately by year as shown in @fig-polarFreqYear, which shows that most of the time the wind is from a south-westerly direction with wind speeds most commonly between 2--6 m s^-1^. In 2000 there seemed to be a lot of conditions where the wind was from the south-west (leading to high pollutant concentrations at this location). The data for 2003 also stand out due to the relatively large number of occasions where the wind was from the east. Note the default colour scale, which has had a square-root transform applied, is used to provide a better visual distribution of the data.



```{r}
#| label: fig-polarFreqYear
#| fig-cap: Use of `polarFreq` function to plot wind speed/directions by year. Each cell gives the total number of hours the wind was from that wind speed/direction in a particular year. The number of hours is coded as a colour scale shown to the right. The scale itself is non-linear to help show the overall distribution. The dashed circular grey lines show the wind speed scale. The date range covered by the data is shown in the strip.
#| fig-width: 8
#| fig-height: 8
polarFreq(mydata, type = "year")
```

The `polarFreq` function can also usefully consider pollutant concentrations. @fig-polarFreqMean shows the mean concentration of SO~2~ by wind speed and wind direction and clearly highlights that SO~2~ concentrations tend to be highest for easterly winds and for 1998 in particular.

By weighting the concentrations by the frequency of occasions the wind is from a certain direction and has a certain speed, gives a better indication of the conditions that dominate the overall mean concentrations. @fig-polarFreqWeightedMean shows the weighted mean concentration of SO~2~ and highlights that annual mean concentrations are dominated by south-westerly winds i.e. contributions from the road itself --- and not by the fewer higher hours of concentrations when the wind is easterly. However, 2003 looks interesting because for that year, significant contributions to the overall mean were due to easterly wind conditions.

These plots when applied to other locations can reveal some useful features about different sources. For example, it may be that the highest concentrations measured only occur infrequently, and the weighted mean plot can help show this.

The code required to make @fig-polarFreqMean and @fig-polarFreqWeightedMean is shown below.

```{r}
#| label: fig-polarFreqMean
#| fig-cap: Use of `polarFreq` function to plot mean SO2 concentrations (ppb) by wind speed/directions and year.
#| fig-width: 8
#| fig-height: 8
polarFreq(mydata, pollutant = "so2", 
          type = "year",
           statistic = "mean", 
          min.bin = 2)
```

```{r}
#| label: fig-polarFreqWeightedMean
#| fig-cap: Use of `polarFreq` function to plot weighted mean SO~2~ concentrations (ppb) by wind speed/directions and year.
#| fig-width: 8
#| fig-height: 8
# weighted mean SO2 concentrations
polarFreq(mydata, pollutant = "so2", 
          type = "year",
           statistic = "weighted.mean", 
          min.bin = 2)
```

Users are encouraged to try out other plot options. However, one potentially useful plot is to select a few specific years of user interest. For example, what if you just wanted to compare two years e.g. 2000 and 2003? This is easy to do by sending a subset of data to the function. Use here can be made of the [openair]{.pkg} `selectByDate` function (see @sec-selectByDate).

```{r}
#| eval: false
# wind rose for just 2000 and 2003
polarFreq(selectByDate(mydata, year = c(2000, 2003)),
          cols = "turbo",
          type = "year")
```

The `polarFreq` function can also be used to gain an idea about the wind directions that contribute most to the overall mean concentrations. As already shown, use of the option `statistic =   "weighted.mean"` will show the percentage contribution by wind direction and wind speed bin. However, often it is unnecessary to consider different wind speed intervals. To make the plot more effective, a few options are set as shown in @fig-polarFreqAttr. First, the `statistic = "weighted.mean"` is chosen to ensure that the plot shows concentrations weighted by their frequency of occurrence of wind direction. For this plot, we are mostly interested in just the contribution by wind direction and not wind speed. Setting the `ws.int` to be above the maximum wind speed in the data set ensures that all data are shown in one interval. Rather than having a square-root transform applied to the colour scale, we choose to have a linear scale by setting `trans = FALSE`. Finally, to show a 'disk', the `offset` is set at 80. Increasing the value of the offset will narrow the disk.

While @fig-polarFreqAttr is useful --- e.g. it clearly shows that concentrations of NO~x~ at this site are totally dominated by south-westerly winds, the use of `pollutionRose` for this type of plot is more effective, as shown in @sec-windRose.

```{r}
#| label: fig-polarFreqAttr
#| fig-cap: The percentage contribution to overall mean concentrations of NO~x~ at Marylebone Road.
#| fig-width: 5
#| fig-height: 5
#| out-width: 60%
polarFreq(mydata, pollutant = "nox", ws.int = 30, 
          statistic = "weighted.mean",
          offset = 80, trans = FALSE, 
          col = "heat")
```
